---
title: "Bellabeat Casestudy"
author: "Alexander Gandji"
date: '2022-06-06'
output:
  html_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
---
  

# Introduction

Bellabeat, a high-tech company that manufactures health-focused smart products wants to analyse the usage of one of their products in order to gain insight into how people are already using their smart devices.Then,using this information, she would like high-level recommendations for how these trends can inform Bellabeat marketing strategy.
The main focus of this case is to analyze smart devices fitness data and determine how it could help unlock new growth opportunities for Bellabeat.  

Key Stakeholders:  

+ Primary Stakeholders:  

+  UrškaSršen: Bellabeat’s cofounder and Chief Creative Officer.  

+  SandoMur: Mathematician and Bellabeat’s cofounder; key member of the Bellabeat executive team.  

+ Secondary Stakeholders:  

+  Bellabeat Marketing Analytics Team: A team of data analysts responsible for collecting, analyzing, and reporting data that helps guide Bellabeat’s marketing strategy.

# Ask phase  

## Business Task  

Analyzing data of smart devices to identify current trends, which can be applied to users, to drive Bellabeat’s/our new marketing strategy


## Who are the stake holders  

+ Urška Sršen=> Bellabeat’s cofounder and Chief Creative Officer
+ Sando Mur=> Mathematician and Bellabeat’s cofounder; key member of the Bellabeat executive team
+ Bellabeat marketing analytics team=> A team of data analysts responsible for collecting, analyzing, and reporting data that helps guide Bellabeat’s marketing strategy
  
  
  
  
  
  
# Prepare Phase  

## Where is the data stored?  

The data is stored on Kaggle and provided by Mönius. The data was gathered through a survey via Amazon Mechanical Tusk. The survey has 33 eligible participants.


## Data limitation  

We are dealing with a small sample size of only 33 participants, which is not a representative sample size. Furthermore, the data set does not provide any information regarding the following aspects;  

+ Social background of the participants (education, profession etc.)
+ When the participants were active, what activity were they doing?
+ Where was the study performed?  
+ There is no information about the participant's gender
The weight loss data set was not included because only data from 8 participants is available 

### Reliability of the data  

#### Reliability & originality  

The data set is provided by the user Mönius. The user-provided a link to the original source of the data=> https://zenodo.org/record/53894#.X9oeh3Uzaao   

#### Comprehensivness  
This aspect is given because the location of the data set and the link to the data source provide enough information;  

+ Name of the company
+ How many participants participated
+ Definition of each variable  

#### Current  

The data was collected in 2016(in the case study was no data mentioned therefore the assumption was made that the case study is from 2016/2017, which makes the data current)  

#### Cited  

The link to the original source is provided by the Kaggle user Mönius. The authors are mentioned on the website where the original data set is located, the authors are mentioned. On the Kaggle website, it is said who collected the data and in what context were the data collected.  


# Process Phase  

The cleaning process was performed with two tools, Excel Spreadsheet and R

## Excel Spreadsheet  

The spreadsheet was selected because the data sets did not contain an overwhelming amount of data; therefore, the data was sorted and filtered in a spreadsheet to do the initial data screening.
In this process following data values were deleted;  

+ Id=> 1503960366, ActivityDate=> 12.05.2016
+ Id=> 6290855005, ActivityDate=> 10.05.2016
+ Id=> 8253242879, ActivityDate=> 30.04.2016
+ Id=> 8583815059, ActivityDate=> 12.05.2016
These values were removed from the dailyActivity data set because they did not gather any data besides 1440 minutes of sedentary data, which could mean the participants forgot to wear their smart devices that day.
  
  
## R
First, all essential packages for the cleaning and analysis process were installed and added to the library. 
```{r echo=TRUE}
install.packages("tidyverse") 
install.packages("skimr")
install.packages("dplyr")
install.packages("ggplot2")

library(tidyverse)
library(skimr)
library(dplyr)
library(ggplot2)
```
## Upload of files
```{r include=FALSE}
daily_activity <- read_csv("/cloud/project/Bellabeat/Clean_Data/dailyActivity_merged.csv")
daily_intensities <- read_csv("/cloud/project/Bellabeat/Clean_Data/dailyIntensities_merged.csv")
hourly_calories <- read_csv("/cloud/project/Bellabeat/Clean_Data/hourlyCalories_merged.csv")
hourly_intensities <- read_csv("/cloud/project/Bellabeat/Clean_Data/hourlyIntensities_merged.csv")
hourly_steps <- read_csv("/cloud/project/Bellabeat/Clean_Data/hourlySteps_merged.csv")
minute_sleep <- read_csv("/cloud/project/Bellabeat/Clean_Data/minuteSleep_merged.csv")
daily_sleep <- read_csv("/cloud/project/Bellabeat/Clean_Data/sleepDay_merged.csv")
```
## Overview of distinct data values
```{r}
nrow(daily_activity)
nrow(daily_intensities)
nrow(daily_sleep)
nrow(hourly_calories)
nrow(hourly_intensities)
nrow(hourly_steps)
nrow(minute_sleep)
```
to get an overall idea, it is also important to check if all sets (at least the sets whit the same time range like daily, hourly etc.)
```{r}
n_distinct(daily_activity$Id)
n_distinct(daily_intensities$Id)
n_distinct(daily_sleep$Id)
n_distinct(hourly_calories$Id)
n_distinct(hourly_intensities$Id)
n_distinct(hourly_steps$Id)
n_distinct(minute_sleep$Id)
```

## Cleaning the data by dropping all NAs 

Even after the prescreening in the Excel Spreadsheet, the whole process was done again in R by dropping NA values.
```{r}
daily_activity_clean <- daily_activity %>% 
  drop_na()

daily_intensities_clean <- daily_intensities %>% 
  drop_na()

hourly_calories_clean <- hourly_calories %>% 
  drop_na()

hourly_intensities_clean <- hourly_intensities %>% 
  drop_na()

hourly_steps_clean <- hourly_steps %>% 
  drop_na()

minute_sleep_clean <- minute_sleep %>% 
  drop_na()

daily_sleep_clean <- daily_sleep %>% 
  drop_na()
```


# Analysis & Share Process  

The first part of the analysis will give an outline of the participants behavior with descriptive data
```{r warning=FALSE}
total_distance_da <- daily_activity_clean %>% 
  select(TotalDistance)

total_distance_da_ <- as.numeric(unlist(total_distance_da))

hist(total_distance_da_, xlab="Kilometer", labels=TRUE, ylim = c(0,200), main = "Distribution of Total Daily Distance", sub="Frequency of Total Distance of the Participants")
```

The histogram shows us that the majority of the data occurs between 0-10 Kilometers to see if we are right we test that statistically  
```{r}
mode <- function(x){
  u <- unique(x)
  tab <- tabulate(match(x,u))
  u[tab==max(tab)]}

total_distance_da %>% 
  summarize(mean=max(total_distance_da_), median=median(total_distance_da_), mode=mode(total_distance_da_), min=min(total_distance_da_), max=max(total_distance_da_))
```

The calculations show that it is a right-skewed distribution.  
What does that mean for Bellabeat? It gives an indicator of what number of kilometres has been reached the most, which will help for the clustering.  

## Cluster
The histogram gave an indicator of how the data is distributed. Before a cluster is created, it is essential to analyse the already existing level of how the kilometres have been accumulated.  

```{r}
## Sedentary
daily_activity_clean %>% 
  summarize(mean=mean(daily_activity_clean$SedentaryActiveDistance), min=min(daily_activity_clean$SedentaryActiveDistance), max=max(daily_activity_clean$SedentaryActiveDistance))




## LightActivity
daily_activity_clean %>% 
  summarize(mean=mean(daily_activity_clean$LightActiveDistance), min=min(daily_activity_clean$LightActiveDistance), max=max(daily_activity_clean$LightActiveDistance))

##Moderately
daily_activity_clean %>% 
  summarize(mean=mean(daily_activity_clean$ModeratelyActiveDistance), min=min(daily_activity_clean$ModeratelyActiveDistance), max=max(daily_activity_clean$ModeratelyActiveDistance))


##VeryActive
daily_activity_clean %>% 
  summarize(mean=mean(daily_activity_clean$VeryActiveDistance), min=min(daily_activity_clean$VeryActiveDistance), max=max(daily_activity_clean$VeryActiveDistance))
```

Based on the mean, min and max of the different levels, a more suitable cluster for distance/kilometres is created;  

+ rarely activity  
+ lightly activity  
+ moderately activity  
+ very active  

0-3.499 kilometers=> rarely active  

3.5-6.499 kilometers=> lightly active  

6.5-<=10 kilometers=> active  

>=10 kilometers=> very active  

The clustering has been chosen because the histogram indicated that the vast majority of kilometres have been in the range between 0-10 kilometres Therefore everything above can be considered very active.

## Total Steps
```{r warning=FALSE}
total_steps <- daily_activity_clean %>% 
  select(TotalSteps)

total_steps_da <- as.numeric(unlist(total_steps))

hist(total_steps_da, labels = TRUE, xlab = "Steps", ylim = c(0,350), main = "Histogram of Total Steps of Participants", sub="Frequency of Total Steps among Participants")

```

Also, the statistical proof that we are dealing with a right-skewed distribution
```{r}
total_steps %>% 
  summarize(mean=mean(total_steps_da), median=median(total_steps_da), mode=mode(total_steps_da))
```

Knowing that the data of total steps are right-skewed indicates that the majority of steps are on the left side. The mean, min and max will allow for creating a representative cluster:  

+ 0-4.999 steps=> rarely active  

+ 5.000-7.999 steps=> lightly_active  

+ 8.000-12.000 steps=> active  

+ 12.000->= 12.000 steps=> very active

## Calories  

```{r warning=FALSE}
daily_activity_clean %>% 
  summarize(mean=mean(daily_activity_clean$Calories), min=min(daily_activity_clean$Calories), max=max(daily_activity_clean$Calories))

hist(daily_activity_clean$Calories, lables=TRUE, main="Distribution of Calory consumption", xlab = "Calories")
```
There is no info about how many females and males participated; therefore, it is not recommended to make any suggestions on how many calories need to be consumed, but it can be helpful to have this data available when we compare our data (calorie consumption of our customers) with the current data set.

# Hourly breakdown of kilometres, steps and calories  
The following paragraph shows in which hours of the day the highest level of steps, calories and activity levels were archived  
First, it is important to separate the ActicityHour, to make it easier to visualize the calorie consume over a day  

## Calory consumption in the AM period  

```{r}
hours_12am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "12:00:00 AM"))
avg_12am <- mean(hours_12am$Calories)

hours_1am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "1:00:00 AM"))
avg_1am <- mean(hours_1am$Calories)

hours_2am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "2:00:00 AM"))
avg_2am <- mean(hours_2am$Calories)

hours_3am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "3:00:00 AM"))
avg_3am <- mean(hours_3am$Calories)

hours_4am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "4:00:00 AM"))
avg_4am <- mean(hours_4am$Calories)

hours_5am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "5:00:00 AM"))
avg_5am <- mean(hours_5am$Calories)

hours_6am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "6:00:00 AM"))
avg_6am <- mean(hours_6am$Calories)

hours_7am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "7:00:00 AM"))
avg_7am <- mean(hours_7am$Calories)

hours_8am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "8:00:00 AM"))
avg_8am <- mean(hours_8am$Calories)

hours_9am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "9:00:00 AM"))
avg_9am <- mean(hours_9am$Calories)

hours_10am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "10:00:00 AM"))
avg_10am <- mean(hours_10am$Calories)

hours_11am <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "11:00:00 AM"))
avg_11am <- mean(hours_11am$Calories)

time_am <- c("12:00:00", "1:00:00", "2:00:00", "3:00:00", "4:00:00", "5:00:00", "6:00:00", "7:00:00", "8:00:00", "9:00:00", "10:00:00", "11:00:00")
avg_am <- c(avg_12am, avg_1am, avg_2am, avg_3am, avg_4am, avg_5am, avg_6am, avg_7am, avg_8am, avg_9am, avg_10am, avg_11am)
avg_am <- sort(avg_am)
hours_am <- data.frame(time_am, avg_am)

hours_am
```

The table shows that the peak is at 11 AM. It is also observable that in the time period from 8 AM-11 AM, the participants consumed the most calories in the AM period.  

## Calory consumption in the PM period  

```{r}
hours_12pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "12:00:00 PM"))
avg_12pm <- mean(hours_12pm$Calories)

hours_1pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "1:00:00 PM"))
avg_1pm <- mean(hours_1pm$Calories)

hours_2pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "2:00:00 PM"))
avg_2pm <- mean(hours_2pm$Calories)

hours_3pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "3:00:00 PM"))
avg_3pm <- mean(hours_3pm$Calories)

hours_4pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "4:00:00 PM"))
avg_4pm <- mean(hours_4pm$Calories)

hours_5pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "5:00:00 PM"))
avg_5pm <- mean(hours_5pm$Calories)

hours_6pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "6:00:00 PM"))
avg_6pm <- mean(hours_6pm$Calories)

hours_7pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "7:00:00 PM"))
avg_7pm <- mean(hours_7pm$Calories)

hours_8pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "8:00:00 PM"))
avg_8pm <- mean(hours_8pm$Calories)

hours_9pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "9:00:00 PM"))
avg_9pm <- mean(hours_9pm$Calories)

hours_10pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "10:00:00 PM"))
avg_10pm <- mean(hours_10pm$Calories)

hours_11pm <- hourly_calories_clean %>% 
  filter(str_detect(ActivityHour, "11:00:00 PM"))
avg_11pm <- mean(hours_11pm$Calories)

time_pm <- c("12:00:00", "01:00:00", "02:00:00", "03:00:00", "04:00:00", "05:00:00", "06:00:00", "07:00:00", "08:00:00", "09:00:00", "10:00:00", "11:00:00")
avg_pm <- c(avg_12pm, avg_1pm, avg_2pm, avg_3pm, avg_4pm, avg_5pm, avg_6pm, avg_7pm, avg_8pm, avg_9pm, avg_10pm, avg_11pm)

hours_pm <- data.frame(time_pm, avg_pm)

hours_pm

```

To better understanding what these numbers mean, following graphs have been generated, which display the calorie consumption behavior in the AM and PM period.  


```{r warning=FALSE}
##AM
calories_plot_am <- ggplot(hours_am, aes(x=time_am, y=avg_am))+
  geom_bar(stat = "identity")+
  labs(title = "Avg. Calory Consumption in the AM Period")+
  theme(axis.text = element_text(angle = 45))

calories_plot_am+
  scale_x_discrete(limits =time_am)

##PM
calories_plot_pm <- ggplot(hours_pm, aes(x=time_pm, y=avg_pm))+
  geom_bar(stat = "identity")+
  labs(title = "Avg Calory Consumption PM")+
  theme(axis.text = element_text(angle = 45))

calories_plot_pm+
  scale_x_discrete(limits =time_pm)
```


The graphs show that the calorie consumption is constantly increasing from 4 AM to 11 AM, which reaches its maximum at 11 AM. In the PM period it is observable at around 12 PM, 2 PM and 5 PM-9 PM that the participants consumed the most calories. Next, the steps and distance walked will be observed.  


## Steps accumulated in the AM period  

```{r}
steps_12am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "12:00:00 AM"))
step_avg_12am <- mean(steps_12am$StepTotal)

steps_1am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "1:00:00 AM"))
step_avg_1am <- mean(steps_1am$StepTotal)

steps_2am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "2:00:00 AM"))
step_avg_2am <- mean(steps_2am$StepTotal)

steps_3am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "3:00:00 AM"))
step_avg_3am <- mean(steps_3am$StepTotal)

steps_4am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "4:00:00 AM"))
step_avg_4am <- mean(steps_4am$StepTotal)

steps_5am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "5:00:00 AM"))
step_avg_5am <- mean(steps_5am$StepTotal)

steps_6am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "6:00:00 AM"))
step_avg_6am <- mean(steps_6am$StepTotal)

steps_7am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "7:00:00 AM"))
step_avg_7am <- mean(steps_7am$StepTotal)

steps_8am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "8:00:00 AM"))
step_avg_8am <- mean(steps_8am$StepTotal)

steps_9am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "9:00:00 AM"))
step_avg_9am <- mean(steps_9am$StepTotal)

steps_10am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "10:00:00 AM"))
step_avg_10am <- mean(steps_10am$StepTotal)

steps_11am <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "11:00:00 AM"))
step_avg_11am <- mean(steps_11am$StepTotal)

steps_hours_am <- c("12:00:00 AM", "1:00:00 AM", "2:00:00 AM", "3:00:00 AM", "4:00:00 AM", "5:00:00 AM", "6:00:00 AM", "7:00:00 AM", "8:00:00 AM", "9:00:00 AM", "10:00:00 AM", "11:00:00 AM")
steps_avg_am <- c(step_avg_12am, step_avg_1am, step_avg_2am, step_avg_3am, step_avg_4am, step_avg_5am, step_avg_6am, step_avg_7am, step_avg_8am, step_avg_9am, step_avg_10am, step_avg_11am)

steps_am_df <- data.frame(steps_hours_am, steps_avg_am)

steps_am_df
```

The people tend to be more active in the time period from 7 AM-11 AM. An odd observation is at 2 AM, with a relatively high average. This can be due to the small sample size.  

## Steps accumulated in the PM period  

```{r}
steps_12pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "12:00:00 PM"))
step_avg_12pm <- mean(steps_12pm$StepTotal)

steps_1pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "1:00:00 PM"))
step_avg_1pm <- mean(steps_1pm$StepTotal)

steps_2pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "2:00:00 PM"))
step_avg_2pm <- mean(steps_2pm$StepTotal)

steps_3pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "3:00:00 PM"))
step_avg_3pm <- mean(steps_3pm$StepTotal)

steps_4pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "4:00:00 PM"))
step_avg_4pm <- mean(steps_4pm$StepTotal)

steps_5pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "5:00:00 PM"))
step_avg_5pm <- mean(steps_5pm$StepTotal)

steps_6pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "6:00:00 PM"))
step_avg_6pm <- mean(steps_6pm$StepTotal)

steps_7pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "7:00:00 PM"))
step_avg_7pm <- mean(steps_7pm$StepTotal)

steps_8pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "8:00:00 PM"))
step_avg_8pm <- mean(steps_8pm$StepTotal)

steps_9pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "9:00:00 PM"))
step_avg_9pm <- mean(steps_9pm$StepTotal)

steps_10pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "10:00:00 PM"))
step_avg_10pm <- mean(steps_10pm$StepTotal)

steps_11pm <- hourly_steps_clean %>% 
  filter(str_detect(ActivityHour, "11:00:00 PM"))
step_avg_11pm <- mean(steps_11pm$StepTotal)

steps_hours_pm <- c("12:00:00 PM", "1:00:00 PM", "2:00:00 PM", "3:00:00 PM", "4:00:00 PM", "5:00:00 PM", "6:00:00 PM", "7:00:00 PM", "8:00:00 PM", "9:00:00 PM", "10:00:00 PM", "11:00:00 PM")
steps_avg_pm <- c(step_avg_12pm, step_avg_1pm, step_avg_2pm, step_avg_3pm, step_avg_4pm, step_avg_5pm, step_avg_6pm, step_avg_7pm, step_avg_8pm, step_avg_9pm, step_avg_10pm, step_avg_11pm)

steps_pm_df <- data.frame(steps_hours_pm, steps_avg_pm)

steps_pm_df
```

The most steps in the PM period are accumulated at 12PM, 2Pm and 5PM-7PM-  
These findings are presented in the following graphs.  

```{r warning=FALSE}
steps_plot_am <- ggplot(steps_am_df, aes(x=steps_hours_am, y=steps_avg_am))+
  geom_bar(stat = "identity")+
  labs(title = "Avg. Steps in the AM Period")+
  theme(axis.text = element_text(angle = 45))

steps_plot_am+
  scale_x_discrete(limits =steps_hours_am)

steps_plot_pm <- ggplot(steps_pm_df, aes(x=steps_hours_pm, y=steps_avg_pm))+
  geom_bar(stat = "identity")+
  labs(title = "Avg. Steps in the PM Period")+
  theme(axis.text = element_text(angle = 45))

steps_plot_pm+
  scale_x_discrete(limits =steps_hours_pm)
```
It is a familiar behavior observable, people tend to be more active around 12 PM and 5 PM-7 PM. There could be a trend noticeable, which will be analyzed later on.

## Intensities  
The data set "hourly_intensities_clean" provides information about how many minutes in each hour were spent doing activities  
```{r}
hourly_intensities_clean

mean(hourly_intensities_clean$TotalIntensity)
```

The data indicates that the participants spent 12 min on average in every hour of the day nonetheless, it would be helpful to understand how the active minutes are distributed over the day.

## Distribution of Active Minutes during the AM period  

```{r}
inten_12am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "12:00:00 AM"))
inten_avg_12am <- mean(inten_12am$TotalIntensity)

inten_1am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "1:00:00 AM"))
inten_avg_1am <- mean(inten_1am$TotalIntensity)

inten_2am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "2:00:00 AM"))
inten_avg_2am <- mean(inten_2am$TotalIntensity)

inten_3am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "3:00:00 AM"))
inten_avg_3am <- mean(inten_3am$TotalIntensity)

inten_4am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "4:00:00 AM"))
inten_avg_4am <- mean(inten_4am$TotalIntensity)

inten_5am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "5:00:00 AM"))
inten_avg_5am <- mean(inten_5am$TotalIntensity)

inten_6am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "6:00:00 AM"))
inten_avg_6am <- mean(inten_6am$TotalIntensity)

inten_7am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "7:00:00 AM"))
inten_avg_7am <- mean(inten_7am$TotalIntensity)

inten_8am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "8:00:00 AM"))
inten_avg_8am <- mean(inten_8am$TotalIntensity)

inten_9am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "9:00:00 AM"))
inten_avg_9am <- mean(inten_9am$TotalIntensity)

inten_10am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "10:00:00 AM"))
inten_avg_10am <- mean(inten_10am$TotalIntensity)

inten_11am <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "11:00:00 AM"))
inten_avg_11am <- mean(inten_11am$TotalIntensity)

inten_hours_am <- c("12:00:00 AM", "1:00:00 AM", "2:00:00 AM", "3:00:00 AM", "4:00:00 AM", "5:00:00 AM", "6:00:00 AM", "7:00:00 AM", "8:00:00 AM", "9:00:00 AM", "10:00:00 AM", "11:00:00 AM")
inten_avg_am <- c(inten_avg_12am, inten_avg_1am, inten_avg_2am, inten_avg_3am, inten_avg_4am, inten_avg_5am, inten_avg_6am, inten_avg_7am, inten_avg_8am, inten_avg_9am, inten_avg_10am, inten_avg_11am)

inten_am_df <- data.frame(inten_hours_am, inten_avg_am)

inten_am_df
```

The table shows that the most active hours during the AM period is the slot between 8 AM-11 AM.  

## PM  

```{r}
inten_12pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "12:00:00 PM"))
inten_avg_12pm <- mean(inten_12pm$TotalIntensity)

inten_1pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "1:00:00 PM"))
inten_avg_1pm <- mean(inten_1pm$TotalIntensity)

inten_2pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "2:00:00 PM"))
inten_avg_2pm <- mean(inten_2pm$TotalIntensity)

inten_3pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "3:00:00 PM"))
inten_avg_3pm <- mean(inten_3pm$TotalIntensity)

inten_4pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "4:00:00 PM"))
inten_avg_4pm <- mean(inten_4pm$TotalIntensity)

inten_5pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "5:00:00 PM"))
inten_avg_5pm <- mean(inten_5pm$TotalIntensity)

inten_6pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "6:00:00 PM"))
inten_avg_6pm <- mean(inten_6pm$TotalIntensity)

inten_7pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "7:00:00 PM"))
inten_avg_7pm <- mean(inten_7pm$TotalIntensity)

inten_8pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "8:00:00 PM"))
inten_avg_8pm <- mean(inten_8pm$TotalIntensity)

inten_9pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "9:00:00 PM"))
inten_avg_9pm <- mean(inten_9pm$TotalIntensity)

inten_10pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "10:00:00 PM"))
inten_avg_10pm <- mean(inten_10pm$TotalIntensity)

inten_11pm <- hourly_intensities_clean %>% 
  filter(str_detect(ActivityHour, "11:00:00 PM"))
inten_avg_11pm <- mean(inten_11pm$TotalIntensity)

inten_hours_pm <- c("12:00:00 PM", "1:00:00 PM", "2:00:00 PM", "3:00:00 PM", "4:00:00 PM", "5:00:00 PM", "6:00:00 PM", "7:00:00 PM", "8:00:00 PM", "9:00:00 PM", "10:00:00 PM", "11:00:00 PM")
inten_avg_pm <- c(inten_avg_12pm, inten_avg_1pm, inten_avg_2pm, inten_avg_3pm, inten_avg_4pm, inten_avg_5pm, inten_avg_6pm, inten_avg_7pm, inten_avg_8pm, inten_avg_9pm, inten_avg_10pm, inten_avg_11pm)

inten_pm_df <- data.frame(inten_hours_pm, inten_avg_pm)

inten_pm_df
```

By reading the table, it becomes apparent that we are observing the same behavior again. People become the most active around 5 PM-7 PM. 
```{r warning=FALSE}
inten_plot_am <- ggplot(inten_am_df, aes(x=inten_hours_am, y=inten_avg_am))+
  geom_bar(stat = "identity")+
  labs(title = "Avg. Intensities in the AM Period")+
  theme(axis.text = element_text(angle = 45))

inten_plot_am+
  scale_x_discrete(limits =inten_hours_am)

inten_plot_pm <- ggplot(inten_pm_df, aes(x=inten_hours_pm, y=inten_avg_pm))+
  geom_bar(stat = "identity")+
  labs(title = "Avg. Intensities in the PM Period")+
  theme(axis.text = element_text(angle = 45))

inten_plot_pm+
  scale_x_discrete(limits =inten_hours_pm)
```

The most active minutes of the day are accumulated during noon and the time period of 5 PM-7 PM. These specific periods have already been observed with the variables. 

## Sleep  

Next, it is also important to get an idea of what the sleeping behavior of the participants is
```{r}
daily_sleep_clean

avg_time_asleep <- round((mean(daily_sleep_clean$TotalMinutesAsleep)/60))
avg_time_in_bed <- round((mean(daily_sleep_clean$TotalTimeInBed)/60))

avg_time_asleep
avg_time_in_bed
```

The average participant slept 7h a day and spent 8h in bed, an assumption would be that participants need 30 min to fall asleep, and it takes them 30 min to get up and leave the bed.  

### Breaking down sleep, calories, activity, steps and distance on weekdays  

The best way to do so is to merge the dailyActivity_clean and sleepDay_clean.  

```{r}
act_sleep_df <- merge(x=daily_activity_clean, y=daily_sleep_clean, c("Id"))
head(act_sleep_df)
```

#### Transforming into weekdays  

```{r}
date_act_sleep_df <- act_sleep_df %>% 
  rename(date=ActivityDate) %>% 
  mutate(date=as.Date(date, format="%m/%d/%Y"))

weekday_act_sleep_df <- date_act_sleep_df %>% 
  mutate(weekday=weekdays(date))

weekday_act_sleep_df$weekday <- ordered(weekday_act_sleep_df$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

weekday_act_sleep_df <- weekday_act_sleep_df %>% 
  group_by(weekday) %>% 
  summarize(wd_steps=mean(TotalSteps), wd_sleep=mean(TotalMinutesAsleep), wd_calories=mean(Calories), wd_distance=mean(TotalDistance))

weekday_act_sleep_df
```

#### Breaking it down for each 

```{r warning=FALSE}
ggplot(weekday_act_sleep_df, aes(x=weekday, y=wd_steps))+geom_bar(stat = "identity")+labs(title="Average Steps during a Week")
```  
  
  
  
  
People tend to walk fewer steps on Sundays. The peak days are Monday, Tuesday and Saturday.  

```{r warning=FALSE}
ggplot(weekday_act_sleep_df, aes(x=weekday, y=wd_sleep))+geom_bar(stat = "identity")+labs(title="Average Minutes Asleep during a Week")
```

Regarding sleep, the sleeping pattern did not change much over the period of a week (avg. 7h).  

```{r warning=FALSE}
ggplot(weekday_act_sleep_df, aes(x=weekday, y=wd_calories))+geom_bar(stat = "identity")+labs(title="Average Calory consumption during a Week")
```

People tend to consume fewer calories on Tuesdays, Wednesdays and Sundays.  

```{r warning=FALSE}
ggplot(weekday_act_sleep_df, aes(x=weekday, y=wd_distance))+geom_bar(stat = "identity")+labs(title="Average Walked Distance during a Week")
```

People walk less on Sundays. The most distance walked is on Tuesdays.

# Correlation & R-squared
After analyzing the descriptive data and getting a good impression of how a typical smart device user might behave, the following paragraph will analyze observable trends of the data. The correlation and R-squared will do this. A quick review/explanation of what the correlation and R-squared are. The correlation allows analysts to determine if there is a strong relationship between two values. R-squared is used to determine how good a value (independent variable) can be used to predict a dependent variable.  

## Calories as dependent variable  
### Total Distance vs Calories  

```{r}
dist_cal <- daily_activity_clean %>% 
  select(TotalDistance, Calories)

cor(dist_cal)
```

The outcome is a correlation of 0.6427, a moderate correlation level (Mukaka, 2012).  

To calculate R-squared it is required to create the function rsq, to calculate R-squared.  
```{r}
rsq <- function (x) cor(x)^2

rsq(dist_cal)
```

R-squared is 0.413, a low predictor for the dependent variable calories (Fernando, 2021).  

```{r warning=FALSE}
ggplot(dist_cal, aes(x=TotalDistance, y=Calories))+
  geom_point()+geom_smooth()+
  labs(title = "TotalDistance vs Calories")+
  geom_label(label= "R^2=0.413", x=17, y=2900, fill=NA)
```

### Total Steps vs Calories  

```{r warning=FALSE}
tost_cal <- daily_activity_clean %>% 
  select(TotalSteps, Calories)

cor(tost_cal)

rsq(tost_cal)

ggplot(tost_cal, aes(x=TotalSteps, y=Calories))+
  geom_point()+geom_smooth()+
  labs(title = "TotalSteps vs Calories")+
  geom_label(label= "R^2=0.344", x=21000, y=2200, fill=NA)
```

The outcome shows that TotalSteps and Calories do not have a high correlation, and TotalSteps is not a good preceptor for Calories (Mukaka, 2012), (Fernando, 2021).  

#### Different levels of activities vs Calories  

Next, the correlation between intensity and calories will be measured. It is important to remember that activity is divided into different levels; each level will be analysed.  

##### Sedentary Minutes vs Calories  

```{r warning=FALSE}
in_s_cal <- daily_activity_clean %>% 
  select(SedentaryMinutes, Calories)

cor(in_s_cal)
rsq(in_s_cal)

ggplot(in_s_cal, aes(x=SedentaryMinutes, y=Calories))+
  geom_point()+geom_smooth()+
  labs(title = "SedentaryMinutes vs Calories")+
  geom_label(label= "R^2=0.008", x=1000, y=500, fill=NA)
```

The conclusion is that sedentary minutes do not have a strong correlation with calories and sedentary minutes are a bad predictor of the calorie consumption.  

##### Lightly Active Minutes vs Calories  

```{r warning=FALSE}
in_la_cal <- daily_activity_clean %>% 
  select(LightlyActiveMinutes, Calories)

cor(in_la_cal)
rsq(in_la_cal)

ggplot(in_la_cal, aes(x=LightlyActiveMinutes, y=Calories))+
  geom_point()+geom_smooth()+
  labs(title = "Lightly Active Minutes vs Calories")+
  geom_label(label= "R^2=0.073", x=300, y=1000, fill=NA)
```

It is a stronger predictor than sedentary minutes but does not correlate strongly with calories.  

##### Fairly Active Minutes vs Calories  

```{r warning=FALSE}
in_fa_cal <- daily_activity_clean %>% 
  select(FairlyActiveMinutes, Calories)

cor(in_fa_cal)
rsq(in_fa_cal)

ggplot(in_fa_cal, aes(x=FairlyActiveMinutes, y=Calories))+
  geom_point()+geom_smooth()+
  labs(title = "FairlyActiveMinutes vs Calories")+
  geom_label(label= "R^2=0.087", x=75, y=1000, fill=NA)
```

The strongest correlation so far, but once again not a strong predictor.  

##### VeryActiveMinutes vs Calories  

```{r warning=FALSE}
in_va_cal <- daily_activity_clean %>% 
  select(VeryActiveMinutes, Calories)

cor(in_va_cal)
rsq(in_va_cal)

ggplot(in_va_cal, aes(x=VeryActiveMinutes, y=Calories))+
  geom_point()+geom_smooth()+
  labs(title = "VeryActiveMinutes vs Calories")+
  geom_label(label= "R^2=0.386", x=100, y=2000, fill=NA)
```

Very Active Minutes do have the strongest correlation and are the best-suited predictor out of all the activity minutes.  

# Key Findings  

+ Based on the data set we have created a new clustering  
  + Kilometers  
    + 0-3.499 kilometers=> rarely active  
    + 3.5-6.499 kilometers=> lightly active  
    + 6.5->=10 kilometers=> active  
    + 10->=10 kilometers=> very active  
    + 0-4.999 steps=> rarely active  
  + Steps  
    + 0-4.999 steps=> rarely active  
    + 5.000-7.999 steps=> lightly_active  
    + 8.000-12.000 steps=> active  
    + 12.000->= 12.000 steps=> very active  

+ People consume the most calories at 12 PM and the time period of 5 PM-7 PM  
+ People walk the most steps at 12 PM and the time period between 5 PM-7 PM  
+ The most active minutes during the day are accumulated during 12pm and the time period of 5pm-7pm  
+ People are the most active on Mondays and Tuesdays, the least on Sundays  
+ Total distance and calories do have a strong relationship  
+ Total Steps and calories do have a strong relationship  
+ Different levels of activities and calories do have a good relationship  
  + The activity level "Very active" has the strongest relationship with calories  


# Act Phase  

The analysis has gained significant insights into how people use their smart devices. The following paragraph will use the findings and outline possible actions for the Bellabeat App.  
The Bellabeat app provides users with their health data related to their activity, steps, habits etc. it allows users to understand their current habits and guides them to healthy decisions.  
The app allows us to apply our gained knowledge the best. Therefore we make the following suggestions;  

+ Sending reminder for food consumption  
  + Calorie consumption based on activity levels
    As we already have established, activity levels do have an acceptable correlation with calories.      The higher the activity level, the bigger the correlation gets. Based on that knowledge, the            app should send a reminder for every bigger meal like breakfast, lunch and dinner. Additional it      should send around 11:30 a notification of how many calories have been already consumed the same       before 5pm (5pm-7pm the most activity time) and a reminder after the workout at 7pm.  
    
+ Calorie consumption based on distance and steps  
The used cluster can here be applied; every time a new cluster level has reached, a notification should be sent out if the user has already eaten, as already established total distance and total steps do have a reasonable correlation with calories.  

+ Giving a weekly overview for users  
It can be helpful to let, every user see their statistic over a week, which will give them a sense of accountability. Based on that, the app can provide suggestions where they need to improve, more steps, more active minutes etc.  
This can be measured first on the analyzed data and eventually will be substituted with primary data, which will be gathered through the app.  

+ Motivating users to move  
A clustering was already introduced. People should reach every day at least the active cluster(8.000-12.000 steps, 6.5-10.000 kilometres). We know the people are very active on Mondays and Tuesdays but do the least on Sundays. They understand that the app should notify people when they seem not to reach the active status. This can be done by showing how many steps kilometres are missing to the goal of letting them know they have reached their goals on previous days and should not let go.  


## Further/Future Analysis  
In the future, we should focus more on our target group, women. The data set, which was analyzed, gave us a great insight about smart device users, but it did not mention how many women participated in the data collection. Moving forward, we should compare our data to the data set and look for similarities or different trends.  
Collecting primary data/data of our users can be obtained through surveys and analyzing the data gathered through the app.


# Appendix
### Sleep vs Calories  

To analyze the sleep and calory variables the most efficient way, it is the recommended approach to merge both data sets.  

#### Merging the datasets & Calculation
```{r}
act_sleep_df <- merge(x=daily_activity_clean, y=daily_sleep_clean, c("Id"))
head(act_sleep_df)

act_sleep <- act_sleep_df %>% 
  select(TotalMinutesAsleep, Calories)

cor(act_sleep)

rsq(act_sleep)
```

## Sleep as a Dependent Variable
### Total Distance vs Sleep  

```{r}
td_sleep <- act_sleep_df %>% 
  select(TotalDistance, TotalMinutesAsleep)

cor(td_sleep)
rsq(td_sleep)
```

Total distance does not have a strong relationship with the sleeping time of the participants; it instead goes the opposite way.   

### Intensity levels vs Sleep
#### Sedentary Minutes vs Sleep  

```{r}
int_s_sleep <- act_sleep_df %>% 
  select(SedentaryMinutes, TotalMinutesAsleep)

cor(int_s_sleep)
rsq(int_s_sleep)
```

#### Lightly Active Minutes vs Sleep  

```{r}
int_la_sleep <- act_sleep_df %>% 
  select(LightlyActiveMinutes, TotalMinutesAsleep)

cor(int_la_sleep)
rsq(int_la_sleep)
```

#### Fairly Active Minutes vs Sleep  

```{r}
int_fa_sleep <- act_sleep_df %>% 
  select(FairlyActiveMinutes, TotalMinutesAsleep)

cor(int_fa_sleep)
rsq(int_fa_sleep)
```

#### Very Active Minutes vs Sleep  

```{r}
int_va_sleep <- act_sleep_df %>% 
  select(VeryActiveMinutes, TotalMinutesAsleep)

cor(int_va_sleep)
rsq(int_va_sleep)
```

#### Total Steps  

```{r}
ts_sleep <- act_sleep_df %>% 
  select(TotalSteps, TotalMinutesAsleep)

cor(ts_sleep)
rsq(ts_sleep)
```


# Sources  

+ Mukaka, M.M. (2012)
  Statistics corner: A guide to appropriate use of correlation coefficient in medical research. Malawi     medical journal : the journal of Medical Association of Malawi, [online] 24(3), pp.69–71. Available     at: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3576830/  
  
+ Fernando, J. (2021)
  R-Squared Definition. [online] Investopedia. Available at:                            https://www.investopedia.com/terms/r/r-squared.asp.  
  
+ Furberg, R., Brinton, J., Keating, M. and Ortiz, A. (2016). Crowd-sourced Fitbit datasets         03.12.2016-05.12.2016. [online] Zenodo. Available at: https://zenodo.org/record/53894#.X9oeh3Uzaao.